---
title: "COMPSCIX 415.2 Homework 3"
author: "Kathleen Welch"
date: "February 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
3.8.1
#1
What is the problem with this plot?
The plot shows mpg in city and hwy but the points don't indicate car class or fuel type so no variable to differenciate the cars.
How could you improve it?
  Adding a color with one of the categories would make the data more meaningful

```{r, fig.align='center',message=FALSE,warning=FALSE}

ggplot(data = mpg, mapping = aes(x = cty, y = hwy, color=class)) + 
  geom_point()+
  xlab("MPG in City")+
  ylab("MPG on Highway")

```

#2
What parameters to geom_jitter() control the amount of jittering?
width and height
```{r}
?geom_jitter()
```

#3
Compare and contrast geom_jitter() with geom_count().
From Description:
Count: Counts the number of observations at each location, then maps the count to point area. 
Jitter: Adds a small amount of random variation to the location of each point.

Compare and contrast 
Both are useful when you have discrete data and overplotting.  Jitter randomly moved the points to stop overlapping - count counts the points ar a given coordinate and reflects that in the size of the point.  So for large datasets of discrete data where the point is exact it is useful - if the points are close but not exact Jitter is more useful for visualization.

```{r}
?geom_count()
?mpg
```
#4
What’s the default position adjustment for geom_boxplot()? 
position_dodge


Create a visualisation of the mpg dataset that demonstrates it.

## No variable
```{r}
ggplot(data = mpg, mapping = aes(x=class,y=cty,fill=as.factor(cyl))) +
geom_boxplot()
```
Using position=dodge looks same

```{r}
?geom_boxplot()
ggplot(data = mpg, mapping = aes(x=class, y=cty,fill=as.factor(cyl))) +
geom_boxplot(position='dodge')
```
Using position jittel looks different...  BTW dodge requires non-overlapping so year worked on x-axis but with warning.  Using position had an error so needed to use a categorical value.
```{r}
ggplot(data = mpg, mapping = aes(x=class, y=cty,fill=as.factor(cyl))) +
geom_boxplot(position='jitter')
```
3.9.1 Exercises


#2    What does labs() do? Read the documentation.
Adds Lables and titles to charts and Graphs

```{r}
?labs()
```

#4 What does the plot below tell you about the relationship between city and highway mpg? 
That both city driving and highway driving improve together.
#Why is coord_fixed() important? 
Makes sure the ratios on the x and y axis are fixed as one - in this case it will give an accurate picture if they were variable the relationship would not be accurate.

#What does geom_abline() do?
Adds a reference line to better annotate the graph

```{r}

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```

Section 4.4: 
#1 Why does this code not work?

my_variable <- 10
my_varıable

Because there is a typo las line needs to be my_variable not my_varıable

corrected
```{r}
my_variable <- 10
my_variable
```


#2 Tweak each of the following R commands so that they run correctly:
Fixed typo 
Fixe fliter to filter and addedd ===
filter to diamonds instead of diamond ....

```{r}
library(tidyverse)
ggplot(data = mpg) +
  geom_point(mapping = aes(x=displ,y=hwy))


filter(mpg, cyl == 8)
filter(diamonds, carat > 3)

```

Section 5.2.4: all exercises. You will need to install the nycflights13 package and use the flights data. 
Hint: to complete exercise #2 you only have to change the code from exercise #1.7.
#1 Find all flights that
        Had an arrival delay of two or more hours
        Flew to Houston (IAH or HOU)
        Were operated by United, American, or Delta
        Departed in summer (July, August, and September)
        Arrived more than two hours late, but didn’t leave late
        Were delayed by at least an hour, but made up over 30 minutes in flight
        Departed between midnight and 6am (inclusive)

    
```{r}
library(nycflights13)

glimpse(flights)

```
```{r}
flights %>% 
 filter(arr_delay > 120)
```
```{r}
flights %>% 
 filter(dest %in% c("IAH", "HOU"))
```

```{r}
flights %>% 
 filter(carrier %in% c("AA","UA","DL"))
```
```{r}
flights %>% 
 filter(month %in% c("7","8","9"))
```
```{r}
flights %>% 
 filter(month, !is.na(dep_delay), dep_delay <= 0, arr_delay > 120)
```
```{r}
flights %>% 
 filter(month, !is.na(dep_delay), dep_delay >= 60, dep_delay-arr_delay > 30)
```
```{r}
flights %>% 
 filter(month, dep_time <=600 | dep_time == 2400)
```
#2 Another useful dplyr filtering helper is between().What does it do? Can you use it to simplify the code needed to answer the previous challenges?

It selects within a list like between 7 and 9 if we change the code from in to between would look like


```{r}
flights %>% 
 filter(month, between(month, 7, 9))
```


#3 How many flights have a missing dep_time? What other variables are missing?
8,225
What might these rows represent? Cancelled flights...
```{r}
flights %>% 
 filter(month, is.na(dep_time))
```
#4 Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)
Not sure what this is asking but if a value is not there is it NA as in missing value.  If there is a value in the field is isn not NA...




5.4.1 Exercises

#1  Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.
Select each field:
select(flights, dep_time, dep_delay, arr_time, arr_delay)
select(flights, starts_with("dep"),starts_with("arr))

```{r}
library(nycflights13)
?flights()
library(tidyverse)

```
#2    What happens if you include the name of a variable multiple times in a select() call?
It removed the duplicate
```{r}
library(nycflights13)

select(flights, dep_time, arr_time, dep_time)
```

#3    What does the one_of() function do? Why might it be helpful in conjunction with this vector?

    vars <- c("year", "month", "day", "dep_delay", "arr_delay")
    
    It allows you to specify the variables within the select command in this case it wold me vars as in:
    select(flights, one_of(vars))
    
    ```{r}
library(nycflights13)
   
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights,  one_of(vars))
```

#4    oes the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

    select(flights, contains("TIME"))
    
    Yes thouht this would be case sensitive - weird that it is not ....
    
    Change to - select(flights, contains("TIME", ignore.case = FALSE))
    
```{r}
library(nycflights13)

select(flights, contains("TIME"))
```

Section 5.4.1: 
all exercises