---
title: "R Notebook"
output: html_notebook
---


Section 3.2.4
#1 1- Run ggplot(data = mpg). What do you see? Nothing
```{r}
ggplot(data = mpg)
```
#2 - How many rows are in mpg? 234  How many columns? 11
---  
mpg
Format

A data frame with 234 rows and 11 variables
----
```{r}
?mpg
```

#3 - What does the drv variable describe? 
  The drive of the car front wheel, rear wheel or 4 wheel.
 
```{r}
?mpg
```
#4 - Make a scatterplot of hwy vs cyl.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(y = cyl, x = hwy))
```
#4 -What happens if you make a scatterplot of class vs drv? it points the type of vechile to the drive - in a plot view.
    Why is the plot not useful?  It is not really useful because there is no relation - a count of drive values to the class would make more sense.
    ```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(y = class, x = drv))
```
Section 3.3.1
#1  -
  What’s gone wrong with this code? 
  gplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
  
  The aesethic color is not a not associated with a variable if you wanted to have colors based on class for example modify syntax to color = class.
  
  Why are the points not blue?
  The points are not blue because if you wanted that to be the case the correct syntax would be 
  gplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
  so that the color doesn’t relate to a variable but is the color to be used for all plotting.
  
#2 -
  Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?
Based on the documentation categorical have fixed values that are drop down - based on the docs that would be drv and class.
```{r}
?mpg
```
  
  Categorical:
  drv
  class
  year
  fl
  cyl
  trans
  
  Continous:
  hwy
  cty
  displ
  

#3
Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?
They show differenct stories and how data is distributed looks different. 
```{r}
?mpg
p <- ggplot(mpg, aes(year,hwy))
p + geom_point(aes(color = drv))
```
#4
What happens if you map the same variable to multiple aesthetics?
Both are applied
```{r}
 ggplot(data = mpg) + 
  geom_point(mapping = aes(x = year, y = hwy, color = drv, shape=drv)) 
```
#5 -
What does stroke do it supports over plotting scatter on numberic value.


```{r}
 ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cyl, y = trans, color=class, stroke = cyl)) 
?geom_point
```
#6
What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?  It evaluates the expression and in this case returns true or false values.
```{r}
 ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cyl, y = trans, color =displ < 5  )) 

```
Section 3.5.1: 
#4
What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?  It breaks the data down into more manageable chunks in this case classifying them by class and putting in 2 rows
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```
#5
Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol argument?
?facet_wrap - splits the data into multiple chunks by to make better use of viewing screen.
nrow - specifies number of rows
ncol - number of columns
Why doesn’t facet_grid() have nrow and ncol argument? need hellp with this one,..

```{r}
?facet_wrap
?facet_grid
?geom_a
```

  
Section 3.6.1: #1-5. 
#1
What geom would you use to draw a line chart? 
geom_line
A boxplot? 
geom_boxplot
A histogram?
geom_hitogram
An area chart?
geom_area

#2
Point map with smooth line 
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE, show.legend = FALSE)
```
#3
What does show.legend = FALSE do? 
It removed the legend message associated with that geom.
What happens if you remove it?
The dash associated with the line is not seen in the legend
Why do you think I used it earlier in the chapter?
It was not needed and could confuse the viewer 

#4
What does the se argument to geom_smooth() do? displays the confidence level (the grey shadow) if it is false does not display.

```{r}
?geom_smooth
```

#5
Will these two graphs look different? Why/why not?
No they should be the same they are different syntax to achieve the same goal.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

#6
first Graph
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

```
Second Graph
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, line=drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

```
3rd Graph
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color=drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

```
4th Graph
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy,)) + 
  geom_point(mapping = aes(color = drv)) + 
  geom_smooth(se = FALSE)

```
5th Graph
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color=drv, linetype=drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

```
6th Graph
```{r}

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color=drv)) + geom_point(shape = 20, color="white", size = 2, stroke = 3)+
geom_point(shape = 20, size = 1, stroke = 2) 

```

3.7.1: 
#2
geom_col uses height values
geom_bar uses proportional data

Question on what is a Data Scientist -
I would not have used the bar chart ... in the first picture stories is confusing it looks like there is only 1% less than available talent unless you read all the stats - if you use  pie chart use colors..
In th holw article it is hard to tell what the story is the use of black and blue doesn't really present a picture of what they are trying to say.