---
title: "Homework 4"
author: "Kathleen Welch"
date: "February 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Go to R for Data Science at http://r4ds.had.co.nz/index.html.

Complete these Exercises:

#Section 5.6.7: #2, #4, #5 and #6 only
#2-a    Come up with another approach that will give you the same output as not_cancelled %>% count(dest) ...

```{r}
library(nycflights13)

not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
not_cancelled %>% count(dest)
```

```{r}

not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
not_cancelled %>% group_by(dest) %>%
  summarize(n = n())

```
 
#2b... and not_cancelled %>% count(tailnum, wt = distance) (without using count()).

```{r}

not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
not_cancelled %>% count(tailnum, wt = distance)


```
```{r}

not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
not_cancelled %>% group_by(tailnum) %>%
summarise(n = sum(distance,na.rm = TRUE))
```
  
#3  Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?
dep_delay - if the fligt is NA then the flight never took off.


#4  Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?

```{r}

cancelled <- flights %>% 
  filter(is.na(dep_delay))
cancelled %>% count(day)
```

```{r}
delayed <- flights %>% 
   filter(!is.na(dep_delay))
delayed  %>% count(day)

```

  #6  Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))

  #7  What does the sort argument to count() do. When might you use it?


#Section 10.5: #1, #2, #3 and #6 only

#Section 12.3.3: #2, #3 and #4 only

#Section 12.4.3: #1 and #2 only


Answer these questions:

Follow these steps:
- Download the baby_names.txt file from Canvas which is in the Homework 4 assignment section.
- Load this file into R correctly and take a glimpse of the output.
- Export this file as an rds file and call it ‘baby_names.rds’.
- Reload the baby_names.rds file and take another glimpse.
- Show all of your code and the output. There should be two data import lines of code, one data export line of code, and two glimpses of the data.
